---
title: "When will He become the GOAT?"
author: "William Foote"
date: '2023-01-09'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(goft)
library(rvest)
```

# Read in the data, combine into one dataframe

```{r}
get_br <- function(url, css_selector) {
  final <- url %>%
    read_html() %>%
    html_element(css = css_selector) %>% 
    html_table()
  final
}
```

```{r, include=FALSE}
rs_2020 <- read.csv('./data/rs_2020.csv')
rs_2020$YR <- '2020'
rs_2021 <- read.csv('./data/rs_2021.csv')
rs_2021$YR <- '2021'
rs_2022 <- read.csv('./data/rs_2022.csv')
rs_2022$YR <- '2022'
rs_2023 <- get_br('https://www.basketball-reference.com/players/j/jamesle01/gamelog/2023',
                  '#pgl_basic > tbody')
names(rs_2023) <- names(rs_2020)[-31]
rs_2023$YR <- '2023'

df <- rbind(rs_2020, rs_2021, rs_2022, rs_2023)
df$Date <- as.Date(df$Date)
df$points_clean <- as.numeric(with(df, ifelse(PTS %in% c('Inactive', 'Did Not Dress', 'Not With Team'), NA, PTS)))
df$game_played_flag <- with(df, ifelse(PTS %in% c('Inactive', 'Did Not Dress', 'Not With Team'), 0, 1))
```

# Analyze Data

## How many games does he miss? How often?

```{r}
# GP = sum(game_played_flag)
sum(df$game_played_flag)
# GP Rate = GP / number of possible games played
sum(df$game_played_flag) / nrow(df)
with(df, plot(game_played_flag ~ (Date)))
```

It looks like He's played about 75% of regular season games in the past three seasons.

## How many points per game does he score when he's playing?

```{r}
# Total PTS in the past 3 seasons
sum(df$points_clean, na.rm = T)
# Overall PPG in past 3 seasons (when He plays)
sum(df[df$game_played_flag == 1, 'points_clean']) / sum(df$game_played_flag)
# By season:
ppg_gp <- df %>% 
  filter(game_played_flag == 1) %>% 
  group_by(YR) %>% 
  summarise(PTS = sum(points_clean, na.rm = T), G = n(), PPG = PTS/G)
ppg_gp
```

## what about when he doesn't play?

```{r}
# PPG in the past 3 seasons (counting missed games as 0's)
ppg_g <- df %>% group_by(YR) %>% 
  summarise(PTS = sum(points_clean, na.rm = T), G = n(), PPG = PTS/G)

ppg_g
```

## Compare the differences of when he plays and doesn't

```{r}
left_join(ppg_gp, ppg_g, by = c('YR')) %>% 
  summarise((G.y * PPG.y) / (G.x * PPG.x))
```

So, we know often he has missed games and how many points he scores when he hasn't miss them. What is the distribution of his scoring patterns? Is it normally distributed, or a little more skewed.

## Get remaining schedule

```{r}
sched <- get_br("https://www.basketball-reference.com/teams/LAL/2023_games.html", css_selector = '#games > tbody')

names(sched)[c(1, 2, 6, 7, 8)] <- c('Game', 'Date', 'Loc', 'Opp', 'Result')

sched <- sched[sched$Date != 'Date', ]

sched$home_flag <- with(sched, ifelse(Loc == '@', 'Away', 'Home'))

```

# Monte Carlo

# he_plays

```{r}
mu_play <- mean(df$game_played_flag)
plays_rng <- sample(c(0, 1), prob = c(1-mu_play, mu_play), size = 1, replace = F)

he_plays <- function() {
  mu_ppg <- mean(df[df$game_played_flag == 1, 'points_clean'], na.rm = T)
  sd_ppg <- sd(df[df$game_played_flag == 1, 'points_clean'], na.rm = T)
  plays_rng <- sample(c(0, 1), prob = c(1-mu_play, mu_play), size = 1, replace = F)
  if (plays_rng == 1) {
    score <- round(rnorm(1, mean = mu_ppg, sd = sd_ppg))
  } else {
    score <- 0
  }
  score
}
```

generates random number to see if he plays,
if he does it also generates how many points he scores

# run_sim

wraps he plays

while score_left > 0, keep running he plays

add this to the game log

```{r}
career_pts <- 37965
career_pts <- get_br(url = 'https://www.basketball-reference.com/players/j/jamesle01.html',
                       '#totals > tfoot')[1, 30][[1]]
run_sim <- function(career_pts) {
  points_left <- 38387 - career_pts
  i <- 1
  while(points_left > 0) {
    this_score <- he_plays()
    if (points_left < this_score) {
      game_log[i, 'points_left'] <- 0
      points_left <- 0
      break
    } else {
      game_log[i, 'points_left'] <- points_left - this_score
      points_left <- points_left - this_score
      i <- i + 1
    }
  }
  i
}
```

```{r}
results <- replicate(10000, run_sim(career_pts))

results_df <- data.frame(game_number = as.integer(names(table(results))), prob = as.integer(table(results))/10000)


fin <- left_join(game_log, results_df, by = 'game_number')
```

### Add exception

if can't find the html, then api limit has been reached
have a backup as an rds