---
title: "normality distribution"
author: "William Foote"
date: '2023-01-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Distribution of PTS scored

### Histogram

A raw look at the data

```{r}
with(df[df$game_played_flag == 1, ], hist(points_clean, prob = T))
```

### Density

To smooth out the distribution

```{r, warning=FALSE}
ggdensity(df$points_clean, fill = 'lightgray')
```
It looks a little bi-modal, with a strong skew on the right side of the graph.

### Check for normality

Is the data normally distributed? Probably not but idk yet

```{r}
ggqqplot(df$points_clean)
```

There are some points that don't lie in the gray area, but not too bad in the right of the graph, actually.

### Run a statistical test for normality

```{r}
df %>% shapiro_test(points_clean)
```

Since the p-value is less than .05, the test implies the data is significantly different from a normal distribution. So, we can't assume normality.

This right-skewed shape kinda looks like a Weibull distribution actually.

## Other tests

### Is it a weibull distribution?

```{r}
weibull_test(df$points_clean)
```

Nope, p-value still tiny.

### What about beta?

Beta distribution is positively (ie all values greater than 0) right skew. Could be applicable here.

```{r}

```
